---
title: "Covid91 vaccine study Final2023"
author: "Ixcell Espinoza- Alvarado"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.



# Question - Does the effectiveness of the vaccine differ based on subpopulation?


# Introduction 

$h_0$ The new preventative vaccine for covid effective for men.

$h_A$ The new preventative vaccine for covid is not effective for men.


# Methods

Variable we use are:

treatment - Categorical

male- Categorical

infected - Categorical

Infected depends on treatment among male

We get just the male from the population

# Results

## Male
 
### Graphical Descriptive Results  
 
```{r}
levels(FinalData$sex)
dataG = filter(FinalData,sex=="male")

barchartGC(~infected + treatment,data=FinalData)
barchartGC(~infected + treatment,data=FinalData, type="percent")

```



By looking at the graph there is a difference between the vaccine and placebo. Looking at the graph the placebo is more effective then the Covid vaccine.


### Numerical discriptive results

```{r}
table1 <- xtabs(~infected + treatment, data=FinalData)
rowPerc(table1)
colPerc(table1)
```


In the top graph, if you were infected with covid19 then 34.99% were getting the vaccine and 65.01% were getting the placebo. If the drug was not effective then the number would be about 50-50. Since the placebo was at a higher percentage it is effective



## Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


With the information above the P-value is 0.000000000000002178 which is <0.05 however we do not fail to reject due to the fact that if the vaccine was not effective then the numbers would be about 50% and 50%. In our case the vaccine was effective therefore we fail to reject the null hypothesis.


## Discussion/Conclusion

IN conclusion the Vaccine is effective on men. And we will fail to reject the hypothesis.





## Females


### Introduction

$h_0$ The new preventative vaccine for covid effective for female.

$h_A$ The new preventative vaccine for covid is not effective for female.

### Methods

Variable we use are:

treatment - Categorical

female- Categorical

infected - Categorical

Infected depends on treatment among male

We get just the male from the population


### Graphical Descriptive Results 

```{r}
levels(FinalData$sex)
dataF = filter(FinalData,sex=="female")

barchartGC(~infected + treatment,data=FinalData)
barchartGC(~infected + treatment,data=FinalData, type="percent")

```


By looking at the graph there is a difference between the vaccine and placebo. Looking at the graph the placebo is more effective then the Covid vaccine.


### Numerical descriptive Results

```{r}
table1 <- xtabs(~infected + treatment, data=FinalData)
rowPerc(table1)
colPerc(table1)
```


In the top graph, if you were infected with covid19 then 34.99% were getting the vaccine and 65.01% were getting the placebo. If the drug was not effective then the number would be about 50-50. Since the placebo was at a higher percentage it is effective.


### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


Similar to the males With the information above the P-value is 0.000000000000002178 which is <0.05 however we do not fail to reject due to the fact that if the vaccine was not effective then the numbers would be about 50% and 50%. In our case the vaccine was effective therefore we fail to reject the null hypothesis.


### Discussion/Conclusion

In conclusion the vaccine is effective on females as well. And we will fail to reject the hypothesis.

## LGBTQ

### Introduction

$h_0$ The new preventative vaccine for covid effective for LGBTQ.

$h_A$ The new preventative vaccine for covid is not effective for LGBTQ.


### Methods

Variable we use are:

treatment - Categorical

LGBTQ- Categorical

infected - Categorical

Infected depends on treatment among male

We get just the male from the population


### Graphical Descriptive Results

```{r}
levels(FinalData$LGBTG)
dataG = filter(FinalData,LGBTQ=="gay")

barchartGC(~infected + treatment,data=dataG)
barchartGC(~infected + treatment,data=dataG, type="percent")

```


By looking at the graph there is a difference between the vaccine and placebo. Looking at the graph the vaccine is more effective then the placebo.



### Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment, data=dataG)
rowPerc(table1)
colPerc(table1)
```


In the top graph, if you were infected with covid19 then 84.62% were getting the vaccine and 15.38% were getting the placebo. If the vaccine was not effective then the number would be about 50-50. Since the vaccine was at a higher percentage it is effective.


### Inferential Descriptive Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


With the information above the P-value is 0.00000006668 which is <0.05 however we do not fail to reject due to the fact that if the vaccine was not effective then the numbers would be about 50% and 50%. In our case the vaccine was effective therefore we fail to reject the null hypothesis.


### Discussion/ Conclusion


In conclusion the vaccine is effective on LGBTQ as well. And we will fail to reject the hypothesis.



## Druggies


# Overall Results and Conclusions