---
title: "Covid91 vaccine study Final2023"
author: "Ixcell Espinoza- Alvarado"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.



# Question - Does the effectiveness of the vaccine differ based on subpopulation?


# Introduction 

$h_0$ There is a difference on effectiveness in the vaccine based on your sex, if you are LGBTQ, or a drug user 

$h_A$ There is not a difference on effectiveness in the vaccine based on sex, if you are LGBTQ, or a drug user 


# Methods

Variable we use are:

treatment - Categorical

male- Categorical

infected - Categorical

Infected depends on treatment among male

We get just the male from the population

# Results

## Male
 
### Graphical Discriptive Results  
 
```{r}
levels(FinalData$sex)
dataG = filter(FinalData,sex=="male")

barchartGC(~infected + treatment,data=FinalData)
barchartGC(~infected + treatment,data=FinalData, type="percent")

```



By looking at the graph there is a difference between the vaccine and placebo. Looking at the graph the placebo is more effective then the Covid vaccine.



## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions